\documentclass[conference]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{cite}
\usepackage{url}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{array}
\usepackage{dblfloatfix}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{placeins}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}


\title{Chain Reaction: LLM-Guided Validation of Kubernetes Attack Chains}

\author{
\IEEEauthorblockN{Ashwin Charathsandran}
\IEEEauthorblockA{
FSCT-8611 Graduation Project\\
British Columbia Institute of Technology (BCIT)\\
Vancouver, BC, Canada
}
}

\begin{document}
\maketitle

\begin{abstract}
Kubernetes security assessments commonly emphasize static configuration scanning and attack-path modeling to prioritize risk. While these techniques identify potentially dangerous conditions, they often fail to demonstrate what is actually exploitable from an in-cluster foothold under runtime constraints. Consequently, defenders may face many plausible multi-step paths with limited evidence indicating which chains can be executed in practice.
\noindent
This project introduces \emph{Chain Reaction}, a Go-based in-cluster agent designed to run as a standard Kubernetes Pod. The agent operates solely with its assigned ServiceAccount credentials and ordinary cluster networking, assuming no node access and no out-of-band secrets. Chain Reaction targets multi-step attack chains spanning RBAC permissions, secret access, network pivots, and takeover preconditions. A chain step is considered validated only when the agent can execute the step from within the Pod using bounded probes and Kubernetes API interactions, and can capture supporting artifacts. Steps that cannot be executed are explicitly classified with failure reasons such as RBAC denial, unreachable network targets, guardrail enforcement, or missing prerequisites.
\noindent
To drive efficient exploration, the system uses an LLM-guided, tool-oriented loop to enumerate relevant cluster objects, summarize effective permissions, test reachability, and validate steps incrementally while enforcing safety guardrails. Guardrails include allow-lists, rate limits, a time budget, and stop conditions to constrain both impact and cost.
\noindent
The primary output is an evidence-backed, phase-labeled attack graph in which each node/edge is annotated with a phase and supported by collected artifacts, paired with an evidence bundle containing raw API responses, probe outputs, timestamps, and object snapshots. Graph edges are labeled as validated or theoretical based on direct runtime evidence. Evaluation is planned in Kubernetes Goat, measuring scenario coverage (targeting validated chains for at least 80\% of scenarios), time-to-chain, API call volume, and run-to-run stability. Results will be compared against theoretical attack-path modeling and in-Pod discovery scanning by reporting the fraction of theoretical paths that become runtime-validated chains.
\end{abstract}

\begin{IEEEkeywords}
Kubernetes, assumed breach, attack-chain validation, RBAC, evidence logging
\end{IEEEkeywords}

\section{Introduction}

\subsection{Background and Motivation}
Kubernetes has become the dominant platform for orchestrating containerized workloads across cloud-native environments. As adoption grows, so does the attack surface: identity-based access through Role-Based Access Control (RBAC), ephemeral workloads, service-to-service networking, and layered security policies all create complex conditions that traditional vulnerability scanners struggle to evaluate. Existing cloud-native security tools focus heavily on static configuration analysis and posture management, producing lists of theoretical attack paths that may or may not be exploitable under actual runtime constraints.

The gap between what \emph{could} happen and what \emph{does} happen when an attacker already has a foothold inside a cluster is significant. Runtime controls such as NetworkPolicies, admission controllers, and restricted Linux capabilities can silently block steps that appear viable on paper. Without a mechanism that tests each step from within the cluster, defenders are left prioritizing risk based on unverified assumptions.

\subsection{Problem Statement}
Current approaches to Kubernetes attack-path analysis identify plausible multi-step chains but do not validate whether those chains are actually executable from an in-cluster identity. Theoretical attack graphs lack runtime evidence, and no existing open-source tool operates from an assumed-breach pod to confirm step-by-step exploitability using bounded, safe probes with auditable artifact collection.

\subsection{Research Questions}
This project is guided by the following research questions:
\begin{enumerate}
    \item \textbf{RQ1:} Can an LLM-guided, tool-calling agent running inside a Kubernetes pod validate multi-step attack chains by executing bounded proof actions against real cluster state?
    \item \textbf{RQ2:} What fraction of theoretically modeled attack paths in a controlled Kubernetes environment (Kubernetes Goat) can be confirmed as runtime-validated chains?
    \item \textbf{RQ3:} How do guardrail mechanisms (allow-lists, rate limits, time budgets, stop conditions) affect the balance between exploration coverage and operational safety?
\end{enumerate}

\section{Literature Review}

Cloud-native security research has changed a lot in the last few years. Instead of only listing possible vulnerabilities, newer work is moving toward tools that can actively test whether an attack is doable in a real environment. This shift matters because Kubernetes clusters are complex: traditional network boundaries are less important, access is mostly controlled through identity, role-based access control (RBAC), and container permissions, and an attack path that looks possible on paper often requires multiple steps. In practice, these chains can fail because of runtime controls such as NetworkPolicies, admission controllers, or missing Linux capabilities.

This literature review summarizes recent work across five themes that relate to validating Kubernetes attack chains. Based on 20 recent academic papers and representative commercial tools, it highlights gaps in current methods and explains why \emph{Chain Reaction} is needed. \emph{Chain Reaction} is designed to run inside a cluster, execute multi-step Kubernetes attack chains from an assumed-breach position, and collect evidence that each step truly worked.

\subsection{Theme 1: LLM-Driven Autonomous Offensive Security}
Large Language Models (LLMs) are increasingly being used to support offensive security. Recent systems show that an agent can plan a multi-step attack, take actions, observe results, and adjust as it goes. However, most evaluations are still based on traditional environments like Linux hosts or web applications. They do not fully cover Kubernetes-specific privilege building blocks such as ServiceAccounts, ConfigMaps, and cluster-scoped RBAC. Table \ref{tab:litmatrix-theme1} summarizes key work in this area.

\begin{table*}[t!]
\centering
\caption{Theme 1: LLM-Driven Autonomous Offensive Security}
\label{tab:litmatrix-theme1}
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l p{2.6cm} X X}
\toprule
\textbf{Author \& Year} & \textbf{Method} & \textbf{Key Findings} & \textbf{Gap vs. Proposal} \\
\midrule
\textbf{Deng et al., 2024} \cite{deng2024pentestgpt} & Hybrid (System Design + Benchmark) & Shows that an LLM can guide multi-step penetration testing using a tool-calling architecture. Supports a basic plan-act-observe loop. & Still depends heavily on a human for execution steps and focuses on standard OS and web targets, not Kubernetes. \\
\addlinespace
\textbf{Fang et al., 2024} \cite{fang2024llm} & Quantitative (Agent eval on 15 CVEs) & Demonstrates that GPT-4 style agents can exploit many 1-day vulnerabilities when given CVE descriptions. & Focuses on single CVEs, not multi-step Kubernetes privilege escalation chains across real cluster configurations. \\
\addlinespace
\textbf{Xu et al., 2024} \cite{xu2024autoattacker} & Hybrid (Multi-agent system design) & Uses separate agents (summarizer, planner, navigator) to manage longer post-breach workflows. & Evaluations assume traditional networks and do not model Kubernetes-native privileges like ServiceAccounts, ConfigMaps, and RBAC. \\
\addlinespace
\textbf{Happe \& Cito, 2023} \cite{happe2023llms} & Quantitative (Priv-esc scenarios) & Benchmarks how well LLMs handle Linux privilege escalation tasks across different scenario difficulties. & Does not capture the distributed nature of Kubernetes, where privilege escalation often involves pivoting between pods, nodes, and APIs. \\
\addlinespace
\textbf{Chen et al., 2024} \cite{chen2024pentestagent} & Hybrid (Multi-agent system) & Splits tasks into recon, exploit, and reporting agents to improve performance over long runs. & Targets enterprise infrastructure more than Kubernetes and does not assume an in-cluster identity model. \\
\addlinespace
\textbf{Happe et al., 2025} \cite{happe2025benchmarking} & Survey and Meta-analysis & Finds that benchmarking in LLM offensive security is often not reproducible across platforms. & Reinforces a gap that \emph{Chain Reaction} addresses by using a standardized baseline environment (Kubernetes Goat). \\
\addlinespace
\textbf{Isamu et al., 2024} \cite{isamu2024towards} & Hybrid (Benchmark Design) & Proposes a standardized benchmark for testing LLM penetration tools across phases like enumeration and exploitation. & Notes the need for more reproducible evaluation setups, including cloud-native contexts, but does not focus on Kubernetes chain execution. \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Theme 2: Attack Graph Modeling and Generation}
Attack graphs are widely used to represent possible lateral movement and privilege escalation paths. Generating these graphs from configuration data is a well-studied problem. The main limitation in cloud-native environments is that the edges in the graph usually represent \emph{possibility}, not confirmed exploitability. In Kubernetes, a graph may indicate an attacker can move from one pod to another, but runtime controls can quietly block the step. What is missing is an execution engine that can test a path from an assumed-breach pod and confirm what is actually reachable. Table \ref{tab:litmatrix-theme2} outlines foundational work in this area.

\begin{table*}[t!]
\centering
\caption{Theme 2: Attack Graph Modeling, Generation, and Analysis}
\label{tab:litmatrix-theme2}
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l p{2.6cm} X X}
\toprule
\textbf{Author \& Year} & \textbf{Method} & \textbf{Key Findings} & \textbf{Gap vs. Proposal} \\
\midrule
\textbf{Ou et al., 2006} \cite{ou2006scalable} & Algorithmic Logic Programming & Uses Datalog-based reasoning over configuration and vulnerabilities to generate scalable attack graphs. & Predates Kubernetes and does not validate edges at runtime. Paths remain theoretical. \\
\addlinespace
\textbf{Ibrahim et al., 2022} \cite{ibrahim2022strategies} & Survey and Hybrid Framework & Surveys modular approaches that separate topology, vulnerability data, and analysis layers to scale attack graphs. & Focuses on enterprise networks and does not provide live execution or proof actions. \\
\addlinespace
\textbf{Bernieri et al., 2023} \cite{bernieri2023scalable} & Modular Framework Design & Proposes an analysis-driven approach that scales to large environments. & Does not account for Kubernetes churn and short-lived workloads. \\
\addlinespace
\textbf{Di Tizio et al., 2024} \cite{ditizio2024coral} & Logical attack graphs with incremental updates & Builds container-focused logical attack graphs that update as pods and containers change. & Models what \emph{could} happen based on RBAC and network state but does not validate reachability through execution. \\
\addlinespace
\textbf{Sultan et al., 2019} \cite{sultan2019generation} & Cloud Topology Retrieval & Studies dynamic attack graph updates by tracking changing IaaS state in cloud providers. & Targets IaaS more than Kubernetes API resources and RBAC-driven permissions. \\
\addlinespace
\textbf{Sultan et al., 2024} \cite{sultan2024detecting} & Systematic Survey & Shows configuration paths often fail at runtime because of policies and missing capabilities. & Supports the need for active validation, but the work is detection-focused rather than execution-focused. \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Theme 3: Container and Cloud-Native Threat Modeling}
Threat modeling work helps define what attackers might target in container systems. Using methods like STRIDE, researchers map threats across images, runtime behavior, orchestration, and networking. These taxonomies are useful because they describe the attack surface that an offensive agent should test. However, most of this work stays at the analysis level. It explains what threats exist but does not show how to autonomously confirm them inside a live cluster with real restrictions. Table \ref{tab:litmatrix-theme3} summarizes this theme.

\begin{table*}[t!]
\centering
\caption{Theme 3: Container and Cloud-Native Threat Modeling}
\label{tab:litmatrix-theme3}
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l p{2.6cm} X X}
\toprule
\textbf{Author \& Year} & \textbf{Method} & \textbf{Key Findings} & \textbf{Gap vs. Proposal} \\
\midrule
\textbf{Getahun et al., 2021} \cite{getahun2021security} & STRIDE Methodology and Attack Analysis & Maps threats across image, runtime, orchestration, and networking layers and provides a clear taxonomy of container attack vectors. & Does not include a method for validating these threats under real cluster constraints. \\
\addlinespace
\textbf{Lin et al., 2025} \cite{lin2025container} & Systematic Survey & Reviews 200+ container vulnerabilities and groups them into many exploit categories. & Strong as a reference, but it does not provide in-cluster execution to test exploitability. \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Theme 4: Adversary Emulation and Breach Simulation Frameworks}
Adversary emulation and Breach and Attack Simulation (BAS) tools are the closest match to what \emph{Chain Reaction} aims to do. Some frameworks also use planning methods to operate in partially observable environments. The main gap is where and how they run. Many systems execute procedures on endpoints (Windows or Linux) or in simulated networks. They usually do not run as a Kubernetes pod using real in-cluster credentials, and they often do not produce step-by-step, auditable evidence artifacts tied to Kubernetes actions. Table \ref{tab:litmatrix-theme4} compares major examples.

\begin{table*}[t!]
\centering
\caption{Theme 4: Adversary Emulation and Breach Simulation}
\label{tab:litmatrix-theme4}
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l p{2.6cm} X X}
\toprule
\textbf{Author \& Year} & \textbf{Method} & \textbf{Key Findings} & \textbf{Gap vs. Proposal} \\
\midrule
\textbf{Miller et al., 2021} \cite{miller2021automated} & AI Planning combined with Caldera & Uses planning approaches to adapt as new environment information is discovered. & Mostly assumes OS-level agents and does not focus on Kubernetes-native steps like ServiceAccount token use. \\
\addlinespace
\textbf{Microsoft Research, 2021} \cite{microsoft2021cyberbattlesim} & Reinforcement Learning Simulation Framework & Provides a platform for training autonomous attacker agents for lateral movement. & Primarily simulated and abstract, not executing real actions against live Kubernetes clusters. \\
\addlinespace
\textbf{Applebaum et al., 2022} \cite{applebaum2022characterizing} & Experimental Evaluation & Finds that Caldera covers many techniques but depends on traditional endpoints. & Does not operate as an in-cluster Kubernetes pod with native credentials and evidence bundles. \\
\addlinespace
\textbf{Lam et al., 2025} \cite{lam2025autonomous} & Multi-agent LLM guided by ATT\&CK & Uses LLM planning to emulate multi-stage adversaries with broad technique coverage. & Evaluated in enterprise network settings and does not use a Kubernetes assumed-breach model. \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Theme 5: Runtime Exploitability Validation}
The purpose of attack chain analysis is to separate theoretical risk from what an attacker can really do. Some research argues that an agent must actively probe its environment to confirm reachability and feasibility under partial observability. The missing piece is applying this idea cleanly inside Kubernetes, where proof actions must work through real API responses, RBAC enforcement, rate limits, and egress limits. Table \ref{tab:litmatrix-theme5} highlights this direction.

\begin{table*}[t!]
\centering
\caption{Theme 5: Runtime Exploitability Validation}
\label{tab:litmatrix-theme5}
\scriptsize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{l p{2.6cm} X X}
\toprule
\textbf{Author \& Year} & \textbf{Method} & \textbf{Key Findings} & \textbf{Gap vs. Proposal} \\
\midrule
\textbf{Garcia et al., 2024} \cite{garcia2024deep} & Deep RL Agent Simulation & Shows that an agent probing its environment can discover attack paths under partial observability. & Results are based on simulation and do not address real Kubernetes constraints like API enforcement, rate limits, or restricted egress. \\
\bottomrule
\end{tabularx}
\end{table*}

\subsection{Competitive Landscape and Synthesis}
Commercial cloud-native security tools often focus on defense and posture management (for example, CNAPP products like Wiz and Orca). Autonomous pentesting platforms (for example, Horizon3.ai's NodeZero and Pentera) are closer to attack-chain execution because they safely run exploits and chain misconfigurations. However, their evidence guarantees, reasoning transparency, and decision logic are mostly proprietary and are not described at an academic level.

Across the five themes, three gaps show up consistently:
\begin{enumerate}
    \item \textbf{Few Kubernetes-native LLM agents:} Existing LLM attack agents usually focus on web apps or OS targets and do not specialize in Kubernetes primitives from an in-cluster identity.
    \item \textbf{Attack graphs stay theoretical:} Many systems can generate plausible paths, but they do not execute bounded proof actions to confirm whether edges are actually traversable.
    \item \textbf{Limited evidence-backed reporting:} Current approaches rarely produce phase-labeled graphs with clear, auditable evidence bundles that prove exploitability step by step.
\end{enumerate}

\emph{Chain Reaction} is designed to address these gaps by moving from theoretical correlation to runtime validation. Instead of assuming a path is reachable, it executes Kubernetes-native steps from an assumed-breach pod and produces artifacts that document what was verified inside the cluster.

\section{Project Approach and Methodology}

\subsection{Overview}
Chain Reaction follows an experimental, execution-based methodology. The system is a Go-based agent deployed as a standard Kubernetes Pod that uses an LLM-guided, tool-calling loop to enumerate cluster objects, test reachability, and validate attack-chain steps incrementally. The agent operates exclusively with its assigned ServiceAccount credentials and ordinary cluster networking, assuming no node-level access and no out-of-band secrets. The project proceeds through four phases.

\subsection{Phase 1: Environment Setup and Baseline}
The evaluation environment is Kubernetes Goat \cite{kubegoat2023}, a deliberately vulnerable Kubernetes cluster that provides a standardized set of misconfigurations and exploitable scenarios. Kubernetes Goat is deployed on a local kind cluster running Kubernetes 1.30. Before any agent runs, each scenario is manually catalogued to establish a ground-truth inventory of expected attack chains and their constituent steps. This catalogue serves as the baseline for measuring scenario coverage.

\subsection{Phase 2: Agent Architecture and Tool Implementation}
The agent is structured around a plan-act-observe loop. In each iteration the LLM receives the current cluster context (discovered objects, effective permissions, prior step outcomes) and selects from a fixed tool set. The core tools include:
\begin{itemize}
    \item \textbf{Kubernetes API Prober} - enumerates namespaces, pods, services, secrets, configmaps, roles, and bindings using \texttt{client-go}.
    \item \textbf{RBAC Analyzer} - resolves effective permissions for the agent's ServiceAccount through \texttt{SelfSubjectAccessReview} and \texttt{SelfSubjectRulesReview} requests.
    \item \textbf{Network Reachability Checker} - tests TCP/HTTP connectivity to discovered service endpoints, the Kubernetes API server, and the kubelet from within the pod.
    \item \textbf{Secret Reader} - attempts retrieval of secret objects to confirm whether RBAC allows access.
    \item \textbf{Token Projector} - inspects mounted ServiceAccount tokens and tests their scope via API authentication checks.
\end{itemize}

All tools are read-only or minimally invasive. No tool creates, modifies, or deletes cluster resources. Safety guardrails constrain the agent's behavior: an allow-list restricts which namespaces and resource types are in scope, a per-step rate limiter caps API requests, a global time budget terminates the run after a configurable duration (default: 5 minutes), and stop conditions halt execution if the agent attempts disallowed actions.

\subsection{Phase 3: Chain Execution and Evidence Collection}
The agent attempts to walk each attack chain step by step. A step is considered \emph{validated} only when the agent can execute the corresponding probe from within the pod and capture a supporting artifact (such as a raw API response body, a TCP handshake result, a decoded secret value, or a permission check output). Steps that cannot be executed are classified with a failure reason: RBAC denial, unreachable network target, missing prerequisite, or guardrail enforcement.

The primary output is a phase-labeled attack graph in which each node represents a cluster object or identity and each edge represents a validated or theoretical transition. Edges are annotated with the execution status and linked to an evidence bundle containing timestamped API responses, probe outputs, and object snapshots.

\subsection{Phase 4: Evaluation and Analysis}
The agent is executed across all Kubernetes Goat scenarios. Each run produces a graph and evidence bundle. Metrics are collected automatically and aggregated across repeated runs. The evaluation compares the agent's validated graph against the manually established baseline and against a static analysis pass that enumerates theoretical paths without runtime execution.

\section{Measurement Models and Metrics}

The evaluation uses four primary metrics:

\begin{enumerate}
    \item \textbf{Scenario Coverage (SC):} The fraction of Kubernetes Goat scenarios for which the agent validates at least one complete attack chain. The target is $\text{SC} \geq 0.80$.
    \begin{equation}
        \text{SC} = \frac{|\{\text{scenarios with} \geq 1 \text{ validated chain}\}|}{|\{\text{total scenarios}\}|}
    \end{equation}

    \item \textbf{Step Validation Rate (SVR):} The fraction of theoretical chain steps (from the baseline) that the agent confirms as executable at runtime.
    \begin{equation}
        \text{SVR} = \frac{|\{\text{validated steps}\}|}{|\{\text{theoretical steps in baseline}\}|}
    \end{equation}

    \item \textbf{Time-to-Chain (TTC):} Wall-clock time in seconds from agent start to the first fully validated chain in a given scenario.

    \item \textbf{API Call Volume (ACV):} Total number of Kubernetes API requests the agent issues per scenario run, used to assess efficiency and cost.
\end{enumerate}

Two secondary metrics supplement the primary set:
\begin{itemize}
    \item \textbf{Run-to-Run Stability:} The coefficient of variation of SC and SVR across five repeated runs per scenario, measuring the consistency of LLM-guided exploration.
    \item \textbf{Guardrail Trigger Rate:} The fraction of runs in which at least one guardrail (rate limit, time budget, allow-list block) was activated, indicating whether safety constraints meaningfully bound the agent's behavior.
\end{itemize}

\section{Statistical and Research Methods}

For each scenario, the agent is executed five times to account for nondeterminism introduced by the LLM's sampling process. Descriptive statistics (mean, standard deviation, coefficient of variation) are computed for SC, SVR, TTC, and ACV across runs.

To assess whether the agent's validated coverage significantly exceeds what static analysis alone produces, a paired Wilcoxon signed-rank test is applied to the per-scenario SVR values for the agent versus the static baseline. This non-parametric test is appropriate given the small sample sizes per scenario and the ordinal nature of step-level validation outcomes. The significance threshold is set at $\alpha = 0.05$.

Correlation analysis (Spearman's rank correlation) is used to examine the relationship between ACV and SVR, testing whether higher API call volume translates to higher validation rates or instead reflects inefficient exploration. Additional analysis examines whether TTC correlates with chain length (number of steps in the baseline chain), which would indicate predictable scaling behavior.

\section{Data Sources and Data Collection}

\subsection{Test Environment}
The primary data source is Kubernetes Goat, an open-source, intentionally vulnerable Kubernetes deployment. It provides a reproducible set of misconfigured scenarios spanning RBAC over-provisioning, exposed secrets, container escapes, and network-level weaknesses. The cluster runs on kind (Kubernetes IN Docker) version 0.20+ with Kubernetes 1.30 on a single-node configuration to ensure consistent state across runs.

\subsection{Data Collected Per Run}
Each agent run produces the following data artifacts:
\begin{itemize}
    \item \textbf{Attack graph:} A JSON-serialized directed graph where nodes are cluster objects or identities and edges carry validation status, phase labels, and evidence references.
    \item \textbf{Evidence bundle:} A directory of timestamped files including raw API responses (JSON), TCP probe results, decoded secret values (redacted where appropriate), RBAC check outputs, and object snapshots.
    \item \textbf{Run log:} A structured log (JSON Lines format) recording each LLM prompt, tool invocation, tool response, decision rationale, and guardrail trigger event with timestamps.
    \item \textbf{Metric summary:} A JSON file recording SC, SVR, TTC, ACV, and guardrail events for the run.
\end{itemize}

\subsection{Volume and Sampling}
Given $N$ scenarios in Kubernetes Goat and 5 runs per scenario, the total dataset comprises $5N$ independent runs. No sampling is applied; all scenarios are included. The manual baseline catalogue is constructed once and frozen before agent evaluation begins.

\section{Data Analysis Techniques}

\subsection{Tools and Software}
Data analysis is performed using Python 3.12 with the following libraries: \texttt{pandas} for tabular aggregation of per-run metrics, \texttt{scipy.stats} for hypothesis testing (Wilcoxon signed-rank, Spearman correlation), \texttt{matplotlib} and \texttt{seaborn} for visualization, and \texttt{networkx} for graph-level analysis of the attack graphs.

\subsection{Analysis Pipeline}
The analysis proceeds in three stages:
\begin{enumerate}
    \item \textbf{Metric extraction:} Run logs and metric summaries are parsed into a unified dataframe indexed by scenario and run number. Attack graphs are loaded into \texttt{networkx} for structural analysis (node count, edge count, connected components, validated-edge fraction).
    \item \textbf{Descriptive analysis:} Per-scenario distributions of SC, SVR, TTC, and ACV are computed and visualized through box plots and bar charts. Heatmaps display per-scenario step validation outcomes across runs.
    \item \textbf{Comparative analysis:} The agent's SVR is compared to the static baseline SVR using a paired Wilcoxon signed-rank test. Spearman correlation tests examine ACV-SVR and TTC-chain-length relationships. Results are reported with effect sizes and confidence intervals where applicable.
\end{enumerate}

\section{Validation and Evaluation Plan}

\subsection{Internal Validity}
To control for LLM nondeterminism, each scenario is run five times and results are reported as distributions rather than single values. The time budget and rate limiter ensure that outlier runs (caused by LLM loops or excessive retries) terminate predictably. The manually constructed baseline provides a fixed reference independent of the agent.

\subsection{External Validity}
Kubernetes Goat is a controlled environment and does not fully represent production clusters. Results should be interpreted as a demonstration of capability within a standardized testbed rather than a claim of general production coverage. The agent's architecture is designed to generalize: tool definitions and allow-lists can be reconfigured for other environments.

\subsection{Baselines and Comparisons}
Two baselines are used for comparison:
\begin{enumerate}
    \item \textbf{Theoretical static analysis:} A configuration-only pass that enumerates all attack paths reachable from the agent's ServiceAccount based on RBAC bindings and network policy definitions, without executing any probes. This represents the current state of the art in attack-graph modeling.
    \item \textbf{In-pod discovery scan:} A single-pass enumeration that queries the Kubernetes API for accessible objects but does not chain steps together or test multi-step reachability. This represents basic in-cluster reconnaissance.
\end{enumerate}

The key evaluation question is how many of the theoretical paths survive runtime validation, and whether chained execution reveals paths or blockers that neither baseline can detect.

\subsection{Reproducibility}
All code, configuration files, deployment manifests, and evaluation scripts are published in the project repository. The Kubernetes Goat environment is deployed from its public Helm chart at a pinned version to ensure reproducibility. Agent configuration (LLM model, temperature, tool definitions, guardrail parameters) is stored in a version-controlled YAML file.

\section{Expected Results}

The anticipated outcomes of this project are as follows.

First, Chain Reaction is expected to achieve a scenario coverage of at least 80\% across Kubernetes Goat scenarios, demonstrating that an in-cluster LLM-guided agent can validate the majority of known attack patterns. Chains involving RBAC over-provisioning and exposed secrets are expected to validate at high rates, while scenarios requiring kernel exploits or container escapes may fall below the validation threshold because the agent operates within standard pod constraints.

Second, the step validation rate is expected to be substantially lower than 100\%, confirming the central hypothesis that many theoretically modeled paths fail at runtime due to NetworkPolicies, missing capabilities, or admission controller enforcement. The gap between theoretical and validated steps is itself a key finding, quantifying the degree to which static analysis overestimates exploitability.

Third, time-to-chain is expected to correlate positively with chain length, with shorter chains (two to three steps) validating in under 60 seconds and longer chains (four or more steps) requiring the full time budget. API call volume is expected to scale roughly linearly with chain length for successful validations.

Fourth, run-to-run stability is expected to show moderate variance (coefficient of variation between 0.10 and 0.30 for SVR), reflecting the inherent nondeterminism of LLM-guided exploration. If variance exceeds this range, it will indicate a need for more structured planning prompts or deterministic fallback strategies.

Finally, the comparison against the static baseline and the discovery scan is expected to demonstrate that runtime validation adds information that neither passive method can provide: specifically, confirming which edges in the theoretical graph are truly traversable and identifying failure reasons for those that are not.

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}